A few stickleback suffice to transport adaptive alleles to new lakes

### 2018-12-12 Stickleback_Paper.tex
plot mod and first attempt at discussion

The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. 
A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations.

@@

\begin{figure}
	\begin{center}
        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome.pdf}
  		\caption{
		Average $F_{ST}$ in windows of 500Mb between:
        		\textbf{(top)} marine habitat and old lakes;
        		\textbf{(middle)} marine habitat and new lakes; and
        		\textbf{(bottom)} old and new lakes.
        		Each plot shows $F_{ST}$ values for a separate simulation,
        		with columns corresponding to increasing gene flow from left to right.
		Ten vertical dotted pink lines in each subplot 
        		show regions which have the potential to affect phenotype.
     } \label{fig:Fst}
	\end{center}
\end{figure}

@@

Now we take a closer look at the genomic architecture of local adaptation between the two habitats. 
%Do the alleles underlying trait differences cluster along the genome?
Do measures of local differentiation identify the causal loci? 
Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between habitats.
(Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
%\plr{need to know where the FAA are in these figures}
Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci 
underlying trait differences between the habitats. 
``Background'' levels of $F_{ST}$ increase as gene flow decreases,
swamping out this signal until the regions under selection are indistinguishable. 
This is likely due to two reasons: 
first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
second, greater sharing of adaptive alleles providing a shared signal across populations.

@@

\plr{Need to fix caption on figure \ref{fig:Power_FP} before editing this bit.}

@@

We have thus far found that the speed of adaptation depends strongly
on the degree to which alleles can be shared between populations.
However, there remains a mystery.
Figure~\ref{fig:MPFAI}A shows that with $m = 0.1$ migrants per lake per generation, 
there is substantial sharing of alleles between populations,
and yet Figure~\ref{fig:TimeToAdaptation} shows that new lakes 
take around 2,000 generations to adapt.
This is surprising because we see a similar quantity of allele sharing at  $m = 1.0$, 
but much more rapid adaptation ($\approx 30$ generations). 
Perhaps
this 60$\times$ discrepancy occurs because 
new lakes must wait for an additional influx of freshwater alleles through migration
beyond what is present in the initial generation?
However, Figure~\ref{fig:Origin} shows that at both $m = 0.1$ and $m = 1.0$,
the vast majority of alleles underlying the local adaptation
derived from individuals in the original generation (``captured'' or ``marine'').

@@

An alternative explanation is that at lower rates of gene flow,
freshwater alleles present as standing variation in the marine habitat
are more tightly linked to marine alleles,
and so adaptation in the new lakes must wait for recombination.
This might be expected since freshwater alleles present in the ocean,
to remain present at migration-selection balance at low migration rates,
must be ``masked'' by nearby compensatory alleles.
Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
Said another way,
higher gene flow from freshwater to the ocean
maintains relatively intact freshwater haplotypes
that can be more easily rebuilt.
Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
in which recombination occurs at rate $10^{-2}$.
Even if there is a sufficient amount of variation in the initial population of a lake
to shift the trait from the marine optimum ($+10$) to the freshwater optimum ($-10$),
these may be tightly linked to alleles that counteract their effect.
For example, suppose there are 10 variants segregating at low frequency with effect size $-1$ each,
but each is paired with a compensatory allele with effect size $+1$.
Each local haplotype is therefore neutral
(and so likely to be at higher frequency in the marine population).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Theoretical expectations}

@@

%sum up the system

%dynamic. and static.

% not subsequent

% reasonable amounts of gene flow result, 
% SGV is powerful, and onlt needs a small amount of gene flow.

% downstream migration not as important. 
% doesn't take a strange environment.

% look up and cite migration load.

We have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum. 
Surprisingly, we have discovered the genetic basis for local adaptation which accelerated by standing genetic variation to be derived largely from the initial generation of inhabitants. 
This is in opposition to the assumption that subsequent migration is also necessary for this process to take place.
In turn, this means that the downstream movement of alleles \emph{from} the lakes \emph{to} the oceans is much more important in this system than upstream migration events. 
We have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
differences between populations that have little gene flow between them.
It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
mistaken for being causative as they often display the same amount of divergence.

\subsection*{thresholds}

@@

\paragraph{Modeling assumptions}
Our simulations had much smaller population sizes than real populations.
How might this affect things?

>>>>

The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. 
What role does genomic variation play in a metapopulation with connection to an array of selective pressures?
How could migration-selection balance actually benefit a super species as a whole in tandem with it experiencing divergent selection?
\jgg{these q's make sense to me, but could be utter non-sense to an experienced pop geneticist. lmk}
A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
In addition, we would like to know how complex geographic systems, with a variety of selective pressures, interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations.

@@

\begin{figure}
	\begin{center}
        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa.pdf}
  		\caption{
		Average $F_{ST}$ in windows of 500Mb between:
        		\textbf{(top)} marine habitat and old lakes;
        		\textbf{(middle)} marine habitat and new lakes; and
        		\textbf{(bottom)} old and new lakes.
        		Each plot shows $F_{ST}$ values for a separate simulation,
        		with columns corresponding to increasing gene flow from left to right.
		Ten vertical dotted pink lines in each subplot 
        		show regions which have the potential to affect phenotype.
		All locations of pre-existing freshwater adapted alleles have been highlighted by 
		blue vertical lines with an alpha value of 0.3, so darker shades of blue imply more 
		freshwater adapted alleles at that location.
     } \label{fig:Fst}
	\end{center}
\end{figure}

@@

Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
Do measures of local differentiation identify the causal loci? 
Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between habitats.
(Note that we are pooling freshwater habitats; a single lake would provide substantially less power.)
\plr{need to know where the FAA are in these figures}
\jgg{check the new plot}
Higher migration rates showed more distinct $F_{ST}$ peaks over polymorphic loci 
underlying trait differences between the habitats. 
``Background'' levels of $F_{ST}$ increase as gene flow decreases,
swamping out this signal until the regions under selection are indistinguishable. 
This is likely due to two reasons: 
first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
second, greater sharing of adaptive alleles providing a shared signal across populations.

@@

\plr{Need to fix caption on figure \ref{fig:Power_FP} before editing this bit.}
\jgg{There is still something wrong with this figure besides the labels ... }

@@

We have thus far found that the speed of adaptation depends strongly
on the degree to which alleles can be shared between populations.
However, there remains a mystery.
Figure~\ref{fig:MPFAI}A shows that with $m = 0.1$ migrants per lake per generation, 
there is substantial sharing of alleles between populations,
and yet Figure~\ref{fig:TimeToAdaptation} shows that new lakes 
take around 2,000 generations to adapt.
This is surprising because we see a similar quantity of allele sharing at  $m = 1.0$, 
but much more rapid adaptation ($\approx 30$ generations). 
Perhaps this 60$\times$ discrepancy occurs because 
new lakes must wait for an additional influx of freshwater alleles through migration
beyond what is present in the initial generation?
This hypothesis is disproven by
Figure~\ref{fig:Origin}, which shows that at both $m = 0.1$ and $m = 1.0$,
the vast majority of alleles underlying the local adaptation
derived from individuals in the original generation (``captured'').

@@

An alternative explanation is that at lower rates of gene flow,
freshwater alleles present as standing variation in the marine habitat
are more tightly linked to marine alleles,
and so adaptation in the new lakes must wait for recombination.
This might be expected since freshwater alleles present in the ocean,
to remain present at migration-selection balance at low migration rates,
must be ``masked'' by nearby compensatory alleles.
Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
Said another way,
higher gene flow from freshwater to the ocean
maintains relatively intact freshwater haplotypes
that can be more easily rebuilt in the marine environments.
Recall that trait-affecting mutations only occur in relatively small regions of $10^5$ loci
in which recombination occurs at rate $10^{-2}$.
Even if there is a sufficient amount of variation in the initial population of a lake
to shift the trait from the marine optimum ($+10$) to the freshwater optimum ($-10$),
these may be tightly linked to alleles that counteract their effect.
For example, suppose there are 10 variants segregating at low frequency with effect size $-1$ each,
but each is paired with a compensatory allele with effect size $+1$.
Each local haplotype is therefore neutral
(and so likely to be at higher frequency in the marine population).

@@

\subsection*{Theoretical expectations}

@@

In this paper, we have described a realistic simulation model to observe the effects of selection of genetic variation (SGV) across a wide range of gene flow rates.
we have shown that historical introgression, at our given parameter sets, is able to reproduce rapid and parallel adaptation similar to what we've seen in real populations such as Middleton island. 
Selection is able to rebuild the freshwater haplotype at a rapid pace (in tens of generations) from marine populations as a medium between all freshwater populations. 
Almost all rates of migration were helpful in the efficiency of the population to locally adapt except for the highest at which migration load limited the ability of the populations to reach the local optimum.

While our findings prove useful to the understanding of stickleback, they are not autonomous to this species. 
Now, we will discuss the broader implications of our findings on the more general framework of similar systems, we also explore what other questions remain. 
Small patches of similar selective pressures can be found in many places such as high altitudes, underwater volcanic vents, or post forest fire environments. 
%We are still finding new subspecies, that seem to have adapted to extreme environments presented in small patches connected to larger spanning biospheres where the ancestral species naturally inhabits.
Recently, (Cite Nelson Ferretti et al), discovered seven new species of Hapalotremus (Theraphosinae), living at the extremely high altitudes distributed along the Andes and Yungas in western South America.
With the ancestral species at much lower altitudes, beneficial high altitude alleles could potentially be carried to help in the colonization of neighboring mountain environments. 
Systems of metapopulations like this could potentially arise in many places
where the ancestral species carries around with it beneficial alleles for
 selective pressures it has encountered in the past and remains connected to through hybridization events.

%(1) the subspeciesthat inhabits a different selective pressure, and ancestral species are not constrained by pre, or post-zygotic isolation to participate in hybridization events, and
%(2) the ancestral species spans across a wide range. 
%For brevity, I will call evolutionary systems which satisfy these conditions, 
%The vast number of species which encou

 \paragraph{Atmosphere's ability to carry whoever scotty is beaming up ... }
Perhaps the most notable finding in this study is that large, ancestral populations have the capacity to maintain alleles which are potentially deleterious for an extended amount of time.  
%This system seems analagous the marvel superhero \emph{Rogue}, who can steal superpowers or memories from another  for a limited amount of time as long as she has contact with the indivi
This principle, deriving solely from the properties of sex and natural selection, could be applicable to almost any species experiencing the influence of divergent selection with introgression.
And to explore further how this process is carried out exactly, many of the questions lie in understanding which properties of genomic architecture and demographics allow this transportation to be possible. 
In terms of Star Trek's ``transporter" referred to by \citet{schluter2009genetics}, 
This is largely analogous to exploring the specific properties of the atmosphere / space which allow for the heroes to be disassembled, then put back together.
%While the perspective of our 
In our simulations, the continued gene flow from freshwater populations meant that the alleles were maintained at a constant rate with fixed demographics. 
Two possible questions that might arise from this are:
(1) How do population sizes, demographics, and evolutionary histories impact an ancestral population's capacity for carrying maladaptive alleles in its opposing environment.
(2) What role does genomic architecture play in the ability for a haplotype to be rebuilt

\paragraph{Origins of genetic basis of local adaptation in new lakes}
In Figure~\ref{fig:Origin} we found that the genetic basis for local adaptation, when selecting upon SGV, is propogated largely from the initial generation of inhabitants. 
This contrasts the belief that subsequent migration is a necessity in newly established populations ability to rebuild a haplotype.
This finding suggests that with high probability, a randomly chosen subset from ancestral species in the metapopulations like marine stickleback, 
carry the capacity to rapidly adapt without \emph{any} continued connection to the rest of of the species.

\paragraph{Realized genomic architecture}
Additionally have also shown introgression is beneficial for inferring causative loci from divergence ($F_{st}$) along the genome. 
This is generally because noise of selectively neutral alleles divergence can appear causative when genetic drift causes more 
differences between populations that have little gene flow between them.
It's important to know that in all scenarios, hitchhiking of selectively neutral alleles could also be 
mistaken for being causative as they often display the same amount of divergence.

%\paragraph{Implications for the ``Transporter"--hypothesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%DO WE NEED THIS?
\begin{comment}
\subsection*{thresholds}

@@

\paragraph{Modeling assumptions}
Our simulations had much smaller population sizes than real populations.
How might this affect things?
\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
